\chapter{Supplementary Evaluation Material}

This appendix should collect material that is useful for reproducibility but too detailed for the main text.

\section{Sample and Task Inventory}

\begin{itemize}
\item \TODO{Add a compact table of the executables used in the final thesis experiments.}
\item \TODO{Summarize the task families attached to each sample, especially the deeper \texttt{config\_decoder\_test} study.}
\item \TODO{Document which samples are stripped, packed, or otherwise designed to stress specific tooling.}
\end{itemize}

\section{Judge and Prompt Artifacts}

\begin{itemize}
\item \TODO{Include abbreviated judge rubric dimensions or a representative rubric excerpt.}
\item \TODO{Include representative task prompts and note how response-scope and analysis-hint wrappers transform them.}
\item \TODO{Add any prompt templates necessary to understand planner, worker, validator, and reporter stages.}
\end{itemize}

\section{Configuration and Experiment Provenance}

\begin{itemize}
\item \TODO{List the main workflow configuration files that define architectures, pipelines, and tool profiles.}
\item \TODO{List the evaluation configuration files that define sweeps, presets, and corpora.}
\item \TODO{Explain how repaired experiments and recovery sessions should be cited in the final artifact package.}
\end{itemize}

\section{Additional Figures or Tables}

\begin{itemize}
\item \TODO{Move large comparison tables, sample manifests, or result-browser screenshots here if they interrupt the narrative in the main text.}
\item \TODO{Add supplementary error-analysis tables if they are too detailed for the main results chapter.}
\end{itemize}
