\chapter{Background and Problem Setting}
\label{sec:background}

\section{Reverse Engineering as an Evidence-Gathering Problem}

As briefly discussed earlier, reverse engineering is not a single question answered by a single tool. Analysts move among disassembly, decompilation, string inspection, structural triage, pattern matching, and artifact correlation in order to construct an evidence-backed explanation of program behavior \cite{mattei2022reusability,yakdan2016johnny}. This makes the domain a strong candidate for studying tool-orchestrated agents.

In practice, reverse-engineering tasks often require both broad survey steps and narrow recovery steps within the same analysis session. Malware-oriented questions also demand conservative claim boundaries, explicit provenance tracking, and visible uncertainty handling. Importantly, the quality of an answer depends on both the final narrative and whether the system gathered the right supporting artifacts along the way.

In the evaluation benchmark, those questions are operationalized as recurring task categories rather than a single generic ``analyze the binary'' prompt. The broadest task family, \emph{default analysis}, asks for program-level behavior reconstruction with evidence anchors and uncertainty handling. Focused tasks then narrow the scope to a single mechanism or artifact family, including configuration or value recovery, decode or parser-flow reconstruction, packing or unpacking triage, and anti-analysis or deception audit. This distinction matters because the evaluation checks not only whether the workflow can produce a plausible narrative but also whether it can pivot from broad overview to targeted recovery when the task demands deeper evidence.

The thesis also uses a concrete notion of acceptable evidence. A strong binary-analysis claim should be tied to identifiable artifacts such as function names or addresses, API or import references, decoded strings or configuration values, registry paths, service names, command lines, or tool outputs such as FLOSS recoveries, capa rules, packer detections, or unpacked-binary metadata. Control-flow explanations are treated as stronger when they connect those artifacts to the branch structure or data path that actually enables the behavior. Reports are penalized when they describe a capability class without showing where the conclusion came from, or when they fail to distinguish observed behavior from inference. That standard later appears directly in the structured judging rubric through dimensions such as \texttt{evidence\_grounding}, \texttt{specificity}, and \texttt{false\_claim\_control}.

\section{LLM Agents, Tool Use, and Orchestration}

The system studied here is explicitly modular. Rather than expecting one model invocation to reason, search, inspect artifacts, and self-critique all at once, the workflow decomposes the task into stages with different responsibilities. This decomposition fits reverse engineering because the task naturally alternates between planning, evidence collection, validation, and synthesis. It also aligns with broader agent work showing that long-horizon performance can benefit from separating high-level planning from low-level execution, that agent-computer interface design can materially affect what the model is able to do in practice, and that iterative critique can improve outputs when review is made explicit \cite{erdogan2025planandactimprovingplanningagents,yang2024sweagentagentcomputerinterfacesenable,madaan2023selfrefineiterativerefinementselffeedback,gou2024criticlargelanguagemodels}.

In this design, preparatory stages gather reusable context, planner stages decompose the query and allocate work across available specialists, and worker stages gather concrete evidence through bounded tool use. Validator stages, when enabled, review coverage, evidence quality, and missing artifacts before the result is finalized. Reporter stages synthesize the resulting findings into the final narrative artifact used by the harness and judge.

\section{MCP-Style Tool Integration}

The system uses MCP-style servers to expose tool capabilities to the agent runtime. This design choice matters because the thesis argument is partly about interface design: how tool affordances are surfaced to agents, how outputs are normalized, and how evidence is routed back into shared state.

As evaluated in this thesis, the operational tool surface is organized around a static-analysis lane plus an artifact-management layer. The static lane provides access to Ghidra and supporting utilities for strings, decoded-string recovery, capability triage, hash resolution, YARA, and binwalk-style file inspection. The artifact layer manages prepared bundles, manifests, run records, judge outputs, and experiment aggregates. The runtime code also reserves metadata for dynamic-tool partitioning, but the active server registry does not currently provide VM, debugger, process-monitor, packet-capture, or other execution-oriented MCP servers.

For the experiments in this thesis, the system is therefore static and artifact-backed. Execution-oriented analysis is not part of the implemented evaluation path, and the stage prompts are written to treat active sample execution as out of scope unless such support exists explicitly.

\section{Binary Analysis Tooling in Scope}

The technical role of each major tool class matters because later architecture chapters build directly on these capabilities.

Ghidra provides decompilation, disassembly, symbol context, and structural inspection, and serves as the primary backend for control-flow and data-flow reasoning. String and FLOSS-style tooling surface literal and decoded text evidence that can anchor behavioral claims. Capability and signature tools provide fast high-level hints that can be corroborated by deeper analysis in later stages. The current implementation does not include an operational dynamic-analysis MCP toolchain, so the evaluated workflow stops at static and artifact-backed evidence rather than live execution or runtime tracing.

\section{Problem Formulation for the Thesis}

At a high level, each experiment asks whether a workflow configuration can answer a binary-analysis task better than a baseline. The atomic evaluation unit is a sample-task pair. That framing carries forward into the evaluation chapter as the basic comparison unit for both authored and held-out experiments.

Each evaluation unit takes as input a binary sample, a task prompt, the available tool set, a workflow configuration, and a model configuration. It produces as output a final analysis report, intermediate evidence artifacts, a judge score, and run metadata. Comparisons are then made along several axes: score deltas, task success rates, error rates, and qualitative failure analysis across matched sample-task pairs.

Seen this way, the thesis is not asking only whether a model can say something sensible about a binary. It is asking three more specific questions: what capability gaps remain once the binaries are obfuscated and artifact-heavy, which workflow choices change performance on those tasks, and whether the final claims are actually backed by the recovered artifacts rather than only by plausible prose.
