\chapter{Related Work}
\label{sec:related}

This chapter positions the thesis against three overlapping research areas: AI-assisted reverse engineering and malware analysis, autonomous or semi-autonomous software agents, and binary-analysis evaluation methodology. The point is not simply to collect related citations, but to clarify what technical gap this thesis addresses.

\section{AI-Assisted Reverse Engineering and Malware Analysis}

Recent academic work addresses several pieces of the design space explored in this thesis. The malware-analysis SoK by Kunwar et al. \cite{kunwar2024sok} shows that transformer-style methods are now applied across static, dynamic, and hybrid malware tasks while emphasizing unresolved problems around obfuscation, explainability, and evaluation. ReSym \cite{xie2024resym} is especially relevant because it demonstrates an analysis-first pattern in which deterministic binary facts are gathered first and then handed to an LLM for higher-level symbol and structure recovery. JTrans \cite{wang2022jtrans}, while focused on binary similarity rather than malware triage, reinforces the broader lesson that control-flow structure matters and that plain token streams are often too weak a representation for difficult binary tasks. Together, these papers justify treating binary analysis as a structured evidence-integration problem rather than a pure prompt-engineering problem.

Benchmark-centric work is also becoming more visible in this area. BinMetric introduces a broad binary-analysis benchmark with six task families and uses it to compare model performance on decompilation-adjacent tasks, summarization, and assembly-oriented outputs \cite{shang2025binmetriccomprehensivebinaryanalysis}. That work provides useful context, but it serves a different purpose: BinMetric is primarily a model-capability benchmark, whereas the present thesis studies a full workflow with staged tool use, shared evidence flow, and configuration-level comparisons over complete sample-task analyses. Very recent contemporaneous work by Hu et al. studies LLM-based pseudocode deobfuscation with a dedicated deobfuscation benchmark and focuses on how reasoning strength, fine-tuning, and obfuscation severity affect deobfuscation quality \cite{hu2026llmsdeobfuscatebinarycode}. That paper provides relevant positioning for obfuscation-heavy reverse engineering, but it is better read as adjacent evidence about deobfuscation-specific model behavior than as a direct precursor to the workflow design studied here.

Recent practitioner-facing systems already point toward the same direction. GhidrAssist \cite{ghidrassist2025}, for example, embeds language-model assistance directly into Ghidra, while GhidraMCP \cite{ghidramcp2025} exposes a structured tool boundary between Ghidra and external agents. These systems are useful comparators because they demonstrate growing interest in LLM-assisted reverse engineering, but they do not by themselves answer the workflow question studied in this thesis: whether planning, specialist decomposition, validation, and experiment-grade measurement materially change performance and robustness within a staged analysis workflow. Raitsis et al. \cite{raitsis2025obfuscation} add a complementary perspective by surveying obfuscation types and detection challenges, which reinforces why malware-facing reverse engineering is a more demanding target than clean binary similarity or source-adjacent code tasks.

\section{Tool-Using Agents and Multi-Agent Orchestration}

The orchestration structure in this thesis is motivated more by the operational needs of reverse engineering than by any claim that one general-purpose agent architecture is already settled in the literature. Reverse engineering naturally decomposes into planning, evidence collection, review, and synthesis, and recent agent systems work suggests that interface design, planning structure, and topology can materially change outcomes. SWE-agent argues that language-model agents benefit from task-specific agent-computer interfaces and shows that interface design can materially affect performance on software engineering tasks \cite{yang2024sweagentagentcomputerinterfacesenable}. Plan-and-Act reports gains from separating high-level planning from low-level execution on long-horizon tasks \cite{erdogan2025planandactimprovingplanningagents}. Multi-Agent Design studies prompt and topology choices jointly and concludes that both contribute materially to multi-agent effectiveness \cite{zhou2026multiagentdesignoptimizingagents}. Abuzakuk et al. further treat workflow optimization as a systems problem, using execution traces to compile recurring tool sequences into deterministic meta-tools that reduce LLM calls and improve robustness on their benchmarks \cite{abuzakuk2026optimizingagenticworkflowsusing}. These works do not study reverse engineering directly, but they support evaluating workflow topology and orchestration structure as real design choices rather than presentation details.

The validator and review aspects of this thesis also overlap with broader critique-and-revision work. Self-Refine studies iterative feedback and refinement using the same model as generator, critic, and refiner \cite{madaan2023selfrefineiterativerefinementselffeedback}. CRITIC extends that idea by incorporating external tools into the critique loop so that revisions can be guided by tool-derived feedback rather than only internal self-evaluation \cite{gou2024criticlargelanguagemodels}. These papers are not reverse-engineering papers, and the thesis implementation uses explicit validator stages rather than generic self-refinement prompting, but they do support treating critique and revision loops as architectural design choices rather than as incidental prompt embellishments.

Prompt-surface changes likewise should not be assumed to help automatically. Basil et al.'s Prompting Science Report 4 finds that expert-persona prompts did not consistently improve factual accuracy on difficult objective QA benchmarks and sometimes degraded it when they were mismatched to the task \cite{basil2025promptingscience4}. That result is not a reverse-engineering result, but it reinforces the thesis decision to treat role-prompt and prompt-shape changes as experimental variables rather than default upgrades.

\section{Program Analysis Toolchains and Human Analyst Workflows}

The human-analyst workflow and interface literature is equally important. Mattei et al. \cite{mattei2022reusability} show that many reverse-engineering tools fail to present results in code context or to support the phases analysts actually move through, while Yakdan et al. \cite{yakdan2016johnny} show that readability-oriented transformations can materially improve analyst performance on malware tasks. At a more structural level, Schwartz et al. \cite{schwartz2018logic} demonstrate the value of explainable program reasoning through rule-based recovery of C++ classes and methods. These works collectively support several design choices in this thesis: keep evidence tied to code context, favor readability and reversible interpretation aids, and preserve deterministic traces or validation logic wherever possible.

The surrounding analyst ecosystem also matters. In addition to decompiler-centric environments, real workflows commonly rely on file- and packer-identification utilities, hash-resolution services, and malware-intelligence feeds. The thesis implementation uses tools in that general mold, including services such as HashDB \cite{hashdb2025}, file-type utilities such as Detect It Easy \cite{detectiteasy2025}, and evaluator-side malware-intelligence references such as MalwareBazaar \cite{malwarebazaar2025}. These tools do not replace structural reverse engineering, but they help shape how analysts prioritize effort, validate hunches, and connect local binary evidence to broader context.

\section{Evaluation of Security and Reverse-Engineering Systems}

The evaluation contribution in this thesis is also narrower than a full benchmark survey. Recent work such as BinMetric \cite{shang2025binmetriccomprehensivebinaryanalysis} is useful because it expands benchmark coverage for binary-analysis question answering and model comparison. The evaluation harness in this thesis serves a different goal: it focuses on sample-task pairs, baseline-first sweeps, repaired experiment provenance, and rubric-driven LLM judging designed for evidence-backed binary-analysis reports. The goal is not to claim a universal evaluation standard for agentic reverse engineering, but to make the tested workflow reproducible enough that prompt-shape, topology, and architecture changes can be compared without losing the underlying task context or artifact trail.


\section{Thesis Positioning}

The distinctive claim of this thesis is not merely that language models can help with reverse engineering. The system contribution is that explicit tool orchestration, staged specialization, and evidence review can be treated as first-class design elements in a reverse-engineering workflow, and that those choices can be evaluated systematically rather than anecdotally.
