\chapter{Implementation and Tool Integration}
\label{sec:implementation}

This chapter translates the high-level workflow into concrete implementation choices. The focus is on how the architecture is realized in software, how tools are exposed to the runtime, and how artifacts and experiment records are preserved. Detailed file-by-file traceability is collected in the appendix so that the main discussion can stay at the system level.

\section{System Organization and Execution Path}

The implementation separates the research system from the measurement apparatus used to study it. The core workflow handles configuration, tool loading, worker topology assembly, staged pipeline execution, and shared-state management. The evaluation infrastructure, described in Chapter~\ref{sec:evaluation}, handles bundles, judging, sweeps, results aggregation, and recovery. This separation matters because it keeps workflow logic independent of the harness mechanics, making it possible to vary one without disturbing the other. Appendix~\ref{app:reproduction-commands} and the supplementary provenance material map these ideas back to the concrete artifact layout for readers who need code-level traceability.

\section{Workflow Configuration}

The workflow uses declarative configuration for stage contracts, worker archetypes, architectures, prompts, and pipeline selection. This design is deliberate: it allows ablations and reproducible experiments to be defined by controlled configuration changes rather than by modifying runtime logic.

Pipeline presets determine which stages run and in what order. Architecture presets determine the available worker mix. Worker prompt shape, role-prompt mode, and review level tune behavior without modifying the core runtime, keeping ablation definitions declarative and auditable.

\section{Ghidra Integration}

Ghidra is the primary structural-analysis backend in the system. What matters here is not only that it is used, but how it is used: through prepared bundles, exported analysis artifacts, and agent-facing access mediated by explicit tooling rather than ad hoc manual interaction.

Ghidra-backed analysis is prepared through bundle generation, which precomputes decompilation and structural metadata before the agent pipeline begins. Agent stages then consume these Ghidra-derived artifacts, and where allowed, they can also query Ghidra-facing tools directly during execution. The system explicitly tracks whether analysis is occurring on the original sample or on a derived artifact such as an unpacked UPX binary so that provenance is preserved across unpacking and re-analysis steps.

At the integration boundary, this work is motivated less by generic chatbot assistance and more by prior evidence that reverse-engineering support tools should improve readability, preserve analyst context, and retain explainable structure \cite{yakdan2016johnny,schwartz2018logic}. Concretely, the system is closest to Ghidra's plugin-mediated extension model \cite{ghidraplugin2025} and to public bridges such as GhidraMCP \cite{ghidramcp2025}. It is also adjacent to assistant-style plugins such as GhidrAssist \cite{ghidrassist2025}, although the emphasis here is not only in-editor assistance. The thesis system uses Ghidra as one component inside a broader staged workflow and evaluation harness, with explicit result capture, replayable bundles, and controlled tool mediation.

\section{Supporting Static Tooling and Artifact Services}

The broader tool ecosystem matters because much of the thesis is about tool augmentation rather than raw model capability.

Static support tools include strings, FLOSS, and hash lookups, while additional utilities provide capability extraction, YARA, binwalk, and UPX-oriented support. Tool profiles allow the evaluation to measure how much accuracy depends on the breadth of this static tool ecosystem by comparing full-access runs against restricted-tool configurations. The runtime architecture reserves metadata for distinguishing static from dynamic tool domains, but the active system registers only static-analysis, enrichment, patching, and artifact-management servers. As a result, the evaluated system does not perform live detonation, debugger orchestration, packet capture, process monitoring, or other execution-driven analysis.

Some of the surrounding tool ecosystem is directly integrated, while other pieces serve mainly as reference points for the analyst workflow the system is trying to emulate. This choice is not only practical; it also follows reverse-engineering usability work showing that analysts benefit from multi-phase workflows, visible evidence transitions, and readable context rather than single opaque outputs \cite{mattei2022reusability,yakdan2016johnny}. HashDB \cite{hashdb2025} is representative of hash-resolution services that help connect encoded imports or constants to likely APIs. Detect It Easy \cite{detectiteasy2025} is representative of file-typing and packer-identification utilities that help scope unpacking or triage effort. MalwareBazaar \cite{malwarebazaar2025} is representative of external malware-intelligence feeds that can contextualize hashes and sample provenance, even though the thesis experiments keep evaluator-side corpus control separate from any live feed dependency.

\section{Artifacts, Bundles, and Provenance}

Tool integration is therefore only one part of the implementation story. The other part is visibility. The runtime records which tools were exposed, which ones were called, how outputs were routed into shared state, and which derived artifacts were retained for later stages or for evaluation. That record is what makes repaired experiments, tool-usage summaries, and failure analysis possible. Without it, the workflow would still produce reports, but it would be much harder to determine whether a weak result reflected poor reasoning, poor tool selection, missing evidence, or an orchestration failure.

Prepared bundles and run artifacts are central to the system's reproducibility story. Bundles capture analysis inputs and precomputed artifacts for a sample, while run records capture the outputs of one particular workflow configuration.

Bundles are reused when their generation inputs remain fresh, and prompt or persona changes do not by themselves invalidate a bundle. Run records preserve final reports, judge outputs, live status, and per-task metadata for each execution. Recovery and repaired-experiment workflows preserve provenance instead of mutating original results in place, ensuring that the distinction between first-pass and recovered runs remains visible in the experiment record.

\section{User-Facing Monitoring and Result Inspection}

The system also includes monitoring and browsing interfaces that operationalize the provenance model and make failure analysis tractable. These interfaces expose in-progress status, archived experiment summaries, per-executable views, and recovery outputs. Their scientific role is not merely convenience: they make it possible to inspect incomplete runs, distinguish analytical weakness from orchestration failure, and preserve the provenance needed for repaired comparisons.
