\chapter{Introduction}
\label{sec:intro}

Software reverse engineering often requires analysts to combine decompilation, string recovery, capability triage, configuration extraction, and other artifact-driven inspection steps. That workflow is tool-rich and evidence-driven, but it is also time-consuming and difficult to automate well. This thesis studies whether an agentic system that explicitly orchestrates multiple tools can produce useful binary understanding on obfuscated malware-oriented binaries, where packing, stripping, encoded artifacts, and anti-analysis behavior complicate the task. The question is increasingly relevant because public reporting and federal complaint data suggest that AI-enabled cyber operations and AI-assisted cybercrime are already appearing outside laboratory demonstrations \cite{anthropic2025espionage,fbi2025ic3}.

The system presented in this thesis is centered on Model Context Protocol (MCP)-style tool orchestration rather than a single monolithic agent. In its maintained form, the workflow combines a light preparatory stage, deterministic presweeps, planning, specialist workers, optional validation, and final reporting, with structured access to Ghidra, supporting static-analysis utilities, and artifact-recovery services. The thesis therefore asks three bounded empirical questions: what capability gaps remain when such a workflow is applied to obfuscated binaries, which workflow variables materially affect analysis performance, and whether the resulting analyses are properly grounded in artifact evidence.

\section{Problem Motivation}

Reverse-engineering tasks are not homogeneous. Some center on control-flow and structural reasoning, while others depend more heavily on recovered strings, decoded configuration data, or packer handling. Because useful evidence is distributed across many tools and intermediate artifacts, and no single model invocation can gather and cross-reference all of it reliably, single-shot prompting is poorly matched to this setting. Tool-augmented agents may improve coverage and evidence quality, but they also introduce orchestration complexity, new failure modes, increased cost, and reproducibility concerns that should be studied rather than assumed.

\section{Research Questions}

The thesis is organized around three primary research questions:
\begin{enumerate}
\item What are the capability gaps in an agent's ability to perform binary analysis on obfuscated binaries?
\item What workflow variables impact performance on analysis capabilities?
\item Are agents able to back claims with proper artifact evidence?
\end{enumerate}

RQ1 focuses on what the implemented system can and cannot recover when binaries are stripped, packed, encoded, or otherwise analysis-resistant. RQ2 studies how performance changes when the workflow configuration changes, using the maintained default workflow as the comparison point for baseline-first sweeps. Here, that comparison point is an \emph{inferred baseline}: the repository's maintained staged workflow retained after development-stage screening, rather than a deliberately weakened strawman. RQ3 then asks whether completed analyses are actually supported by visible artifacts rather than only by plausible narrative.

\section{Thesis Scope and Expected Contributions}

This thesis makes four contributions:
\begin{enumerate}
\item A modular agentic reverse-engineering system that separates preparation, planning, specialist evidence collection, optional review, and reporting so those roles can be varied and measured explicitly.
\item MCP-mediated integration of Ghidra, supporting static-analysis tools, and artifact-recovery services, providing structured tool access rather than ad hoc manual interaction.
\item A binary-analysis evaluation harness built around sample-task pairs, baseline-first sweeps, repeated runs, and rubric-guided LLM-judge scoring, designed to make workflow comparisons reproducible and to preserve evidence-grounding and robustness signals alongside final scores.
\item A focused empirical investigation of three questions: what capability gaps remain on obfuscated binaries, which workflow variables change analysis performance, and when completed analyses are or are not artifact-backed.
\end{enumerate}

\section{Thesis Organization}

The remainder of this thesis is organized as follows. Chapter~\ref{sec:background} introduces the reverse-engineering problem setting and the technical background needed for the thesis. Chapter~\ref{sec:related} positions the work against prior literature on AI-assisted reverse engineering, malware analysis automation, and tool-using agents. Chapter~\ref{sec:overview} describes the system at a high level, and Chapter~\ref{sec:implementation} details the implementation and tool integrations. Chapter~\ref{sec:evaluation} defines the experimental methodology, makes the maintained comparison baseline explicit for RQ2, and maps the experiment artifacts to the three research questions. Chapter~\ref{sec:results} then answers the three questions in order, using archive studies as development-stage context and the repaired held-out evaluation as the main source of evidence. Chapter~\ref{sec:limitations} discusses limitations and threats to validity, and Chapter~\ref{sec:conclusion} closes by restating what the current evidence does and does not support.
