\chapter{Limitations and Threats to Validity}
\label{sec:limitations}

This chapter states the main boundaries on the claims made in this thesis. Some of these limits come from the benchmark itself, some from the model and workflow choices, and some from the measurement and reporting stack. Making those limits explicit is important because the workflow combines nontrivial orchestration, tool integration, structured judging, and repaired experiment outputs.

\section{Benchmark Scope and External Validity}

The final held-out evaluation is stronger than the earlier development studies, but it remains a small benchmark. The thesis-grade comparison covers five held-out executables, one locked default-analysis task per executable, and three repaired repetitions per configuration. That scope is sufficient to demonstrate that executable-level heterogeneity matters, but it is not sufficient to support a broad claim that any single workflow is generally best for malware reverse engineering.

The authored development benchmark introduces a different limitation. It is useful for controlled workflow debugging and ablation design, but it is not a substitute for a large external malware corpus. The archived broad and decoder-depth studies therefore function as methodological evidence and case-study evidence, not as the final basis for generalization.

\section{Statistical Power and Repetition Count}

The repaired final experiment improves coverage, but it does not remove the small-sample problem. Each configuration has only three repaired repetitions, and the final evaluation includes only five executables. That directly limits statistical power. In the final repaired experiment, no overall score comparison reached the thesis significance target of $p \le 0.05$. This means the final chapter can support descriptive comparisons and executable-level case analysis more confidently than strong confirmatory claims about superiority.

The significance outputs should therefore be read as guardrails rather than as the main argument. They help prevent overclaiming when an observed mean difference looks appealing, but they do not overcome the basic limitation that both the replicate count and the executable count remain modest. Additional executables, more repetitions, or both would be needed before making stronger population-level claims.

\section{Model Scope}

The empirical results in this thesis are tied to the specific model budget and workflow choices used in the evaluated runs. In practice, much of the execution used mini- and nano-class models rather than larger frontier models. That matters in at least two ways.

First, some of the observed failure modes are plausibly model-class sensitive. The context-window and request-budget problems that appeared in the final run may behave differently with larger models, different context windows, or different provider-side limits. Second, positive or negative results under mini/nano models do not automatically transfer to larger models. A workflow that looks brittle here may become more robust with a larger context window, while a workflow that looks cost-effective here may lose that advantage if evaluated only with a more capable but more expensive model.

For that reason, the current findings are best read as evidence about \emph{this} workflow under \emph{this} budget-conscious model regime, not as a definitive statement about all agentic reverse-engineering systems.

\section{Recovery and One-Shot Robustness}

The final performance chapter intentionally uses the repaired experiment for score comparison and the unrecovered sweep for robustness analysis. That is the most honest way to preserve both pieces of evidence, but it also introduces a limitation. The repaired experiment is not the same thing as a clean one-shot evaluation in which every task completed successfully on the first attempt.

This matters because recovery is doing real work. In the held-out evaluation, the unrecovered sweep exposed infrastructure-heavy failure modes, especially host-worker context overflows on Ghidra-heavy assignments. Later patches and recovery runs improved completeness materially. That makes the repaired experiment the right artifact for final like-for-like comparison, but it also means that the repaired tables should be interpreted as ``best available completed comparison under the evaluated workflow'' rather than as evidence that the workflow was robust on its first pass.

\section{Tool-Call Redundancy Measurement}

The redundancy reporting added during this work is informative, but it does not justify the strongest possible claims about unnecessary tool use. The current implementation records tool calls per task and preserves a \texttt{source} label that usually maps to a stage manager or worker slot. In other words, the current summaries do \emph{not} first pool together calls from unrelated tasks before counting redundancy. That is enough to distinguish several important cases:

It can distinguish repeated calls within one sample-task run from calls originating in different tasks, same-source repetition from cross-source overlap within the same task, and broad within-task hotspots such as repeated program-info queries or repeated \texttt{todo\_read} state lookups.

However, the current logs still do not attach every tool call to a specific work item or objective inside a reused worker slot. As a result, the reporting can reliably say that a given task and source produced repeated calls, but it cannot always say whether those calls were truly excess. Some repeated calls may represent legitimate cross-stage refreshes, parallel workers collecting shared context within one task, or one worker slot being reused for multiple different objectives inside the same task.

The consequence is important for interpretation. The redundancy figures and tables are good for identifying hotspots, likely same-source loops, and areas where better memory or caching may help. They are \emph{not} strong enough to prove that every repeated call counted in the summaries was wasteful. Stronger instrumentation, especially per-call work-item attribution, would be needed before making a firmer claim about exact avoidable overhead.

\section{Judging and Measurement Proxies}

Judge scores remain a proxy for analyst usefulness, not a direct substitute for expert human review. The rubric is structured and useful, but it still compresses a complicated reverse-engineering outcome into a set of scalar dimensions and one overall score. That creates two related risks.

Some runs may benefit from better organization or cleaner evidence presentation without necessarily achieving deeper mechanistic understanding. Conversely, a partially correct but poorly framed answer may score worse than an analyst would judge it in an interactive workflow where clarification is possible.

The thesis reduces this risk by pairing score reporting with failure-mode reporting, per-executable comparisons, and preserved task artifacts. Even so, the judge should be treated as a consistent evaluation instrument rather than as a perfect oracle.

\section{Operational and Ethical Considerations}

Finally, the workflow is inherently dual-use. Malware-oriented reverse-engineering tooling can assist defensive analysis, but it can also accelerate unsafe understanding of malicious capabilities. The system mitigates that risk through local execution, evaluator-side corpus control, and explicit workflow logging, but those mitigations do not eliminate the dual-use concern. Any broader operational deployment would require stronger containment, access control, and review procedures than those needed for a thesis evaluation environment.
