\chapter{Conclusion and Future Work}
\label{sec:conclusion}

This chapter restates the core problem, summarizes the main findings, and identifies the next research steps.

\section{Conclusion}

This thesis studied whether reverse engineering and malware analysis benefit from a workflow that treats tool use, staged reasoning, review, and result provenance as first-class concerns rather than as incidental prompt-engineering details. The system contribution is a modular agentic workflow that separates preparation, planning, specialist evidence gathering, optional validation, and final reporting, together with an evaluation harness that can run baseline-first sweeps, preserve per-task artifacts, recover failed cases, and regenerate comparison outputs reproducibly.

The empirical contribution is more specific than a claim that one workflow is simply best. The development-stage broad and decoder-depth studies were useful for screening the design space and for identifying which variables were worth carrying into the held-out comparison. They showed that prompt shape and worker topology were more informative than obvious handicaps such as reducing tool access, and they exposed structural failure modes such as validator-heavy collapse before the blind final run was launched.

The repaired held-out evaluation provides the strongest evidence in the thesis. It shows that the system can often produce usable static analyses on held-out stripped, packed, and otherwise obfuscation-heavy binaries, but not with universal reliability or depth. At the same time, no non-baseline configuration achieved a statistically significant overall improvement over the maintained baseline at the thesis confidence target ($p \le 0.05$). The strongest descriptive overall candidate was \texttt{worker\_prompt\_shape:artifact\_focused}, which produced its largest gain on \texttt{sample3.exe} ($+21.4$ over baseline). However, the final comparison also showed strong executable-level heterogeneity: the same configuration could help on one held-out sample and hurt on another, and every non-baseline row was flagged as masked by average in the consistency summary. The clearest evidence-backed takeaway is therefore not that the thesis discovered a universally best configuration, but that capability gaps and workflow-variable effects both depend heavily on the executable being analyzed.

The held-out run also showed that the system's remaining capability gaps are partly operational rather than purely interpretive. The unrecovered sweep exposed context-bloat and request-budget failures that were not dominant in the smaller development benchmarks. Those failures do not erase the repaired comparisons, but they do show that agentic reverse engineering must be evaluated as both a reasoning problem and a workflow-engineering problem. A system that can reason well in principle is still limited if it cannot keep tool outputs, worker contexts, and recovery paths operationally tractable on harder binaries. The efficiency summaries reinforce that point: repeated tool activity is concentrated on a subset of executables and is dominated by specific patterns such as \texttt{todo\_read} loops and selected Ghidra rescans, which suggests that workflow overhead is not purely a model-quality issue.

Taken together, these results answer the three thesis research questions in a bounded way.

\paragraph{RQ1.} The repaired and unrecovered held-out results show that the system can often recover useful static findings on obfuscated binaries, but not with uniform reliability or depth. The main remaining capability gaps are one-shot completion on harder samples, deeper staged or later-stage recovery, and stable management of large or repetitive tool context.

\paragraph{RQ2.} Workflow variables do change performance, but their effects are conditional rather than universal. Relative to the maintained baseline, the strongest descriptive candidate is \texttt{worker\_prompt\_shape:artifact\_focused}, but no tested configuration produces a statistically significant overall gain and every non-baseline variant is mixed across the five held-out executables.

\paragraph{RQ3.} Yes, in a qualified sense. The repaired final rubric summaries show that completed analyses can be evidence-backed, and the evidence-focused prompt variant improves the dimensions most closely tied to grounded reporting. At the same time, this behavior is not guaranteed: baseline evidence grounding is only moderate on average, and some configurations degrade false-claim control or specificity.

These answers reposition the main contribution of the thesis. The work does not show that one prompt or one topology solves malware-oriented reverse engineering. It shows that agentic reverse engineering should be treated as a capability-gap, workflow-selection, and evidence-grounding problem, and that those three concerns have to be evaluated together.

\section{Future Work}

Several next steps follow directly from these findings.

\paragraph{Broader benchmarks.} Additional held-out executables, more repetitions, and more than one locked task per family would clarify whether the executable-level heterogeneity observed here is a stable pattern or a sample-limited artifact.

\paragraph{Better instrumentation.} Per-call work-item attribution, stronger caching or memory controls, and richer robustness accounting would make it possible to separate analytical weakness from orchestration overhead more cleanly than the current summaries allow. The repaired final run suggests two especially practical targets: stage-level guards around repeated \texttt{todo\_read} loops and better reuse of Ghidra program or import context across preflight, presweeps, and worker stages. Trace-derived workflow compression through meta-tools is one concrete example of the kind of optimization that could reduce repeated tool sequences without changing the underlying model \cite{abuzakuk2026optimizingagenticworkflowsusing}.

\paragraph{Model-class generalization.} The system should be evaluated beyond the budget-conscious mini/nano regime used in the current runs. Larger models, different provider limits, or different context windows may change both the benefit of staged orchestration and the failure modes that dominate harder samples.

\paragraph{Dynamic analysis.} Execution-oriented analysis would require a more substantive implementation than the current system provides. The architecture reserves metadata for distinguishing static from dynamic tool domains, but the evaluated system does not include a real dynamic-analysis MCP lane. A future study could add controlled execution, runtime tracing, or sandbox evidence collection and then evaluate when that extra evidence genuinely improves grounded analysis rather than only adding complexity.

More broadly, the thesis suggests a shift in emphasis for future research on agentic reverse engineering. The most useful question may not be ``which single workflow is best?'' but rather ``how should a workflow adapt to different executable profiles while remaining grounded, auditable, and operationally reliable?'' That is a harder problem than prompt tuning alone, but it is also the more realistic one for reverse engineering as a long-term analyst-assistance direction.
